Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities

### 2020-03-03 writing/transect_paper.Rmd
finished first pass through coauthor comments

We attempted to denoise both long and short PacBio amplicons using `DADA2` according to the steps outlined in the supplementary information in @callahan2019.
However, only `r rdd(region_table)$ASVs[rdd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
The `DADA2` algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
For the combination of long reads (median $L=`r rdd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(rdd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - rdd(demuxqual)['pb_500'])^rdd(demuxlength)['pb_500'])), big.mark = ",")`.
Given the high diversity relative to sequencing depth in this study (`r rdd(region_table)$ASVs[rdd(region_table)$seq_run == "pb_483" & rdd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r rdd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon ($L=`r rdd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r rdd(demuxqual)['pb_483']`$) is only `r round(2/(1 - rdd(demuxqual)['pb_483'])^rdd(demuxlength)['pb_483'], 1)` reads.
We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:

@@

Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors.
Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using `DADA2` version `r packageVersion("dada2")` [@callahan2016; @callahan2019].
The error model for `DADA2` denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
For PacBio libraries, `DADA2` was run with complete pooling and a band size of 16.
For Ion Torrent libraries, pseudo-pooling and a band size of 32 were used, and the homopolymer gap penalty was set to -1, as recommended by the `DADA2` FAQ (https://benjjneb.github.io/dada2/faq.html).
Chimeras within each region were removed using `removeBimeraDenovoTable` from `DADA2`.

@@

Because the Illumina dataset consisted of paired-end reads, regions were not extracted prior to denoising.
ASVs were instead generated according to a standard workflow for `DADA2`.
Demultiplexed reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
Forward and reverse reads were denoised using `DADA2` version `r packageVersion("dada2")` [@callahan2016] using separate error models and pseudo-pooling, and then forward and reverse reads were merged.
The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the other technologies.

@@

Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@wang2007] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [@adl2019], which utilizes hierarchical nameless ranks.
FASTA format files of the re-annotated RDP and Warcup training sets are available at (*somewhere*).

Taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three algorithms:
the RDP Naïve Bayesian Classifier (RDPC) as implemented in DADA2;
SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
and IDTAXA [@murali2018].
Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

@@

@@

```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
ggpubr::ggarrange(
  rdd(tech_class_pcoa_plot),
  rdd(tech_ecm_fam_pcoa_plot),
  common.legend = TRUE,
  legend = "bottom",
  labels = "auto"
)

# plot(rdd(tech_class_pcoa), type = "none")
# points(
#   rdd(tech_class_pcoa),
#   select = with(
#     rdd(tech_class_sample_data),
#     tech == "PacBio" & amplicon == "Short"
#   ),
#   col = "green",
#   pch = 1
# )
# points(
#   rdd(tech_class_pcoa),
#   select = with(
#     rdd(tech_class_sample_data),
#     tech == "PacBio" & amplicon == "Long"
#   ),
#   col = "blue",
#   pch = 2
# )
# points(
#   rdd(tech_class_pcoa),
#   select = with(
#     rdd(tech_class_sample_data),
#     tech == "Illumina" & amplicon == "Short"
#   ),
#   col = "red",
#   pch = 3
# )
# with(
#   rdd(tech_class_pcoa_scores),
#   text(x = MDS1*2, y = MDS2*2, labels = abbrev)
# )
# legend(
#   "bottomleft",
#   legend = c("Illumina Short", "PacBio Short", "PacBio Long"),
#   col = c("red", "green", "blue"),
#   pch = c(3, 1, 2)
# ) 
# 
# plot(rdd(tech_ecm_fam_pcoa), type = "none")
# points(
#   rdd(tech_ecm_fam_pcoa),
#   select = with(
#     rdd(tech_ecm_fam_sample_data),
#     tech == "PacBio" & amplicon == "Short"
#   ),
#   col = "green",
#   pch = 1
# )
# points(
#   rdd(tech_ecm_fam_pcoa),
#   select = with(
#     rdd(tech_ecm_fam_sample_data),
#     tech == "PacBio" & amplicon == "Long"
#   ),
#   col = "blue",
#   pch = 2
# )
# points(
#   rdd(tech_ecm_fam_pcoa),
#   select = with(
#     rdd(tech_ecm_fam_sample_data),
#     tech == "Illumina" & amplicon == "Short"
#   ),
#   col = "red",
#   pch = 3
# )
# with(
#   rdd(tech_ecm_fam_pcoa_scores),
#   text(x = MDS1*3, y = MDS2*3, labels = abbrev)
# )
# legend(
#   "bottomleft",
#   legend = c("Illumina Short", "PacBio Short", "PacBio Long"),
#   col = c("red", "green", "blue"),
#   pch = c(3, 1, 2)
# )
```

```{r}
tech_class_nmds <- vegan::metaMDS(
  phyloseq::otu_table(rdd(tech_class_table))
)

tech_class_nmds_scores <-
  vegan::scores(tech_class_nmds, display = "species") %>%
  as_tibble(rownames = "class") %>%
  mutate(L = sqrt(NMDS1^2 + NMDS2^2)) %>%
  filter(!is.nan(L)) %>%
  filter(L >= max(L) / 2) %>%
  mutate(abbrev = substr(class, start = 1, stop = 2))

plot(tech_class_nmds, type = "n")
vegan::ordispider(
  tech_class_nmds,
  groups = with(
    rdd(tech_class_sample_data),
    paste(buffer, year, site, x)
    ),
  col = "gray80"
  )
text(
  x = tech_class_nmds_scores$NMDS1,
  y = tech_class_nmds_scores$NMDS2,
  labels = tech_class_nmds_scores$abbrev
  )
points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
                                     tech == "PacBio" & amplicon == "Short" ),
       col = "green", pch = 1)
points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
                                     tech == "PacBio" & amplicon == "Long"),
       col = "blue", pch = 2)
points(tech_class_nmds, select = with(rdd(tech_class_sample_data),
                                     tech == "Illumina" & amplicon == "Short"),
       col = "red", pch = 3)
legend("topright", legend = c("Illumina Short", "PacBio Short", "PacBio Long"), col = c("red", "green", "blue"), pch = c(3, 1, 2))
```

### Positive control

*Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
Additional ASVs belonging to the genus *Agaricus* were also detected in the same wells as the positive control in the Ion Torrent dataset, but these were sufficiently different that they probably represent additional *Agaricus* species from the field samples, and were found in the same wells due to incomplete demultiplexing.
*A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair.
The abundance of the second ASV was `r rdd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~. > 0) %>% percent(format = "f", digits = 2) %>% paste(collapse = " and ")` that of the primary ASV in the two Illumina positive controls.
The consistency of this ratio across PCR replicates suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.

The primary *A. bisporus* ASV was the most abundant ASV in the long-amplicon PacBio dataset and the Ion Torrent dataset, and the fifth most abundant ASV in the short-amplicon PacBio dataset, and the third most abundant ASV in the Illumina dataset.
This ASV was also found in `r rdd(pos_control_data) %>% filter(!is.na(pc_frac)) %>% group_by(seq_run) %>% summarize(well_frac = sum(pc_frac > 0) / n()) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 1) %>% text_list()` of the non-control samples, representing `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively.

@@

Mantel correlograms of the Bray-Curtis and weighted UNIFRAC community dissimilarities with respect to spatial distance and time lag are shown in Figure\ \@ref(fig:correlog).
Results are qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation ($p < 0.05$, filled red circles) in the Bray-Curtis metric at ranges of up to 2--3\ m for the total fungal community, and 1--2\ m for the ECM community.
In both cases, the greatest correlation magnitudes and distances are associated with Illumina sequencing of the short amplicon library, followed by PacBio sequencing of the long amplicon library, with PacBio sequencing of the short amplicon library detecting the least spatial structure.

In contrast, the weighted UNIFRAC distance, which was only calculated for the long amplicon library, shows very little spatial structure, with only the total fungal community in the 1\ m 
distance class showing a significant correlation at $p < 0.05$.

The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation at a time lag of 1\ year and 0\ m separation (filled blue circles) for both the total fungal and ECM communities in the long amplicon library.
For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0\ m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1\ m for both sequencing technologies.
This negative correlation was significant in all short amplicon correlograms.

Distance-decay curves relating community dissimilarities and spatial distances are shown in Figure\ \@ref(fig:spatio).
For the cases where the Mantel correlogram detected a significant positive correlation, the best-fit exponential curve is also displayed; for the long amplicon library, where there was also a significant positive correlation after a time lag of one year, the best fit curve for time lag is also included.
The best fit spatial and temporal ranges, along with a 95% confidence interval, are shown in Table\ \@ref(tab:variofit).
Although a model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m was included in the 95% confidence interval.
Although the best fit spatial ranges vary between the different datasets by a factor of about 3, there is overlap of the 95% confidence intervals for all of the Bray-Curtis variograms across communities, amplicon libraries, and sequencing technologies.

@@

```{r variofit, results="asis"}
rdd(variog_data) %>%
  select(mantel_0, mantel_1, variofit_0 = variofit2, variofit_1 = variofitST2, metric, amplicon, tech, guild, algorithm) %>%
  pivot_longer(1:4, names_to = c(".value", "timelag"), names_sep = "_") %>%
  mutate(params = map(variofit, broom::tidy)) %>%
  unnest(params) %>%
  bind_cols(
    pmap_dfr(
      .,
      function(mantel, variofit, term, ...) {
        if (mantel) {
          tryCatch(broom::confint_tidy(object = variofit, parm = term) %>% as.list %>% inset2("term", term) %>% as.data.frame,
                   error = function(e) tibble(term = term, conf.low = NA_real_, conf.high = NA_real_))
        } else {
          tibble(term = term, conf.low = NA_real_, conf.high = NA_real_)
        }
      }) %>%
      select(-term)
  ) %>%
  filter(term %in% c("range", "timerange")) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

@@

The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 1)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 1)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 1)`),
the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.

DADA2 denoising to form ASVs may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].

Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, no such amplicons were sequenced by any of the technologies, and Illumina showed a slightly higher fraction of longer sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Supplementary Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).

`r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` of long amplicon reads beloged to ASVs which were not found in any short amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences which were much shorter than those recovered from the short amplicon datasets \@ref(fig:ITS2-length).
Taxonomic assignments reveal that many of these non-shared sequences fall outside kingdom Fungi (Supplementary Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Supplementary Figure\ \@ref(fig:short-its)).
Within Fungi, the short amplicon datasets recovered more Sordariomycetes, possibly because this group was represented by diverse but rare taxa which were not detected in the PacBio datasets due to the lower sequencing depth.
Additionally, several groups showed anomalous detection in either the long or short datasets, such as *Tulasnella* in the long amplicon dataset (Supplementary Figure \@ref(fig:ecm-heattree)), and *Xylaria* and *Myerozyma* in the short amplicon datasets (Supplementary Figure\ \@ref(fig:heattree-length-compare)).
These differences may be due to primer mismatches in these groups.
However, neither amplicon recovered any sequences identifiable to *Cantharellus*, an ECM genus which is known from the area from fruitbodies.

@@

ASV recovery for long amplicons was dramatically improved (`r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` to `r percent((rdd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")`) by denoising homologous subregions independently.
Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs, so this approach should increase ASV recovery from these platforms as well.
It may also be adaptable to Oxford Nanopore sequencing, which has hitherto posed difficulties for application to complex community metabarcoding [@loit2019].

@@

The RDP fungal training set and Unite performed comparably at placing long amplicon sequences.
The Warcup database placed notably fewer sequences at all taxonomic levels for all datasets, probably in part due to the fact that only fungal sequences are included.
However, even with this taken into account, IDTAXA performed very poorly with the Warcup database, placing less than 25% of ASVs to kingdom in all datasets.
IDTAXA placed fewer sequences than RDPC or SINTAX even with the other databases, but this is expected given its more conservative assignment of confidence scores [@murali2018a].

@@

For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
For instance, `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "d") %>% rev() %>% text_list()`.
This led to a corresponding increase in the fraction of reads assigned to a functional guild (Supplementary Figure\ \@ref(fig:ecm)).

>>>>

We attempted to denoise both long and short PacBio amplicons using DADA2 according to the steps outlined in the supplementary information in @callahan2019.
However, only `r rdd(region_table)$ASVs[rdd(region_table)$region=="long"]` amplicon sequence variants (ASVs) were obtained for the long amplicons, representing `r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of the trimmed reads.
We conclude that this poor performance was due to a combination of long read length and low sequencing depth relative to community diversity.
The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
For the combination of long reads (median $L=`r rdd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(rdd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - rdd(demuxqual)['pb_500'])^rdd(demuxlength)['pb_500'])), big.mark = ",")`.
Given the high diversity relative to sequencing depth in this study (`r rdd(region_table)$ASVs[rdd(region_table)$seq_run == "pb_483" & rdd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r rdd(demuxcounts)["pb_500"]` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon ($L=`r rdd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r rdd(demuxqual)['pb_483']`$) is only `r round(2/(1 - rdd(demuxqual)['pb_483'])^rdd(demuxlength)['pb_483'], 1)` reads.
We therefore developed a new workflow to assemble ASVs from the long amplicons, as follows:

@@

Each of the extracted regions was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors.
Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")` [R package `dada2`, @callahan2016; @callahan2019].
The error model for DADA2 denoising was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
For PacBio libraries, DADA2 was run with complete pooling and a band size of 16.
For Ion Torrent libraries, pseudo-pooling and a band size of 32 were used, and the homopolymer gap penalty was set to -1, as recommended by the DADA2 FAQ (https://benjjneb.github.io/dada2/faq.html).
Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.

@@

Because the Illumina dataset consisted of paired-end reads, regions were not extracted prior to denoising.
ASVs were instead generated according to a standard workflow for DADA2.
Demultiplexed reads were truncated after the first base with quality score <= 10, and then reads with more than 3 expected errors in either read were discarded.
Forward and reverse reads were denoised using DADA2 version `r packageVersion("dada2")` [@callahan2016] using separate error models and pseudo-pooling, and then forward and reverse reads were merged.
The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the other technologies.

@@

Taxonomic annotations of the Ribosomal Data Project's LSU fungal training set (RDP) version 11.5 [@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to the taxonomic classification system used in the Unite database version 8 [@nilsson2019a].
In particular, the classification for fungi was according to @tedersoo2018, and for non-fungal eukaryotes was according to the proposed system of @tedersoo2017c.
Although the latter system is not formally published, it is consistent with the annotations for non-fungal eukaryotes in the Unite database.
Additionally, it is a system with both purportedly monophyletic taxa and a uniform set of taxon ranks, which make it more appropriate for sequence-based taxonomic assignment algorithms than more accepted classification systems such as that of the International Society of Protistologists [@adl2019], which utilizes hierarchical nameless ranks.
FASTA format files of the re-annotated RDP and Warcup training sets are available at (*somewhere*).

Taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three algorithms:
the RDP Naïve Bayesian Classifier [RDPC, @wang2007] as implemented in DADA2;
SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
and IDTAXA [@murali2018].
Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references $\times$ three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

@@

*Agaricus bisporus*, the positive control, was represented by a single ASV in the positive control samples for both long- and short-amplicon PacBio datasets, and in the Ion Torrent dataset.
*A. bisporus* was represented by two ASVs in the Illumina dataset, which differed at one base pair (`r Biostrings::readDNAStringSet(here::here("data/clusters/ITS2.fasta.gz")) %>% magrittr::extract(startsWith(names(.), "f2ab49e8")) %>% Biostrings::width() %>% unique() %>% divide_by(1, .) %>% subtract(1, .) %>% percent(digits = 1, format = "f")` similarity in ITS2).
The abundance of the second ASV was `r rdd(agaricus_reads) %>% as.data.frame() %>% {.[["f2ab49e8"]]/.[["39d80c19"]]} %>% keep(~. > 0) %>% percent(format = "f", digits = 1) %>% paste(collapse = " and ")` that of the primary *A. bisporus* ASV in the two Illumina positive controls.
The consistency of this ratio across replicate positive controls suggests that it represents true inter-copy variation within the specimen, rather than sequencing or PCR error.
Despite higher total sequencing depth, this ASV was not identified from the Ion Torrent dataset.

*A. bisporus* sequences represented `r rdd(pos_control_data) %>% filter(complete.cases(.)) %>% group_by(seq_run) %>% summarize(read_frac = sum(pc_reads) / sum(all_reads)) %>% deframe() %>% magrittr::extract(c("pb_500", "pb_483", "SH-2257", "is_057")) %>% percent(format = "f", digits = 2) %>% text_list()` of non-control reads, in the PacBio long, PacBio short, Illumina, and Ion Torrent datasets, respectively, giving similar estimates for the rate of tag-switching for all technologies.
These reads were excluded from community analyses.

@@

```{r ppcoa, fig.cap = "(ref:ppcoa-short), (ref:ppcoa)", fig.scap = "(ref:ppcoa-short)"}
ggpubr::ggarrange(
  rdd(tech_class_pcoa_plot),
  rdd(tech_ecm_fam_pcoa_plot),
  common.legend = TRUE,
  legend = "bottom",
  labels = "auto"
)
```

@@

Results of spatial analysis based on the Bray-Curtis dissimilarity were qualitatively similar between the two amplicon libraries and between PacBio and Illumina sequencing, with significant autocorrelation at $p < 0.05$ for ranges of up to 2--3\ m for the total fungal community, and 1--2\ m for the ECM fungal community (Figure\ \@ref(fig:correlog)).
In both cases, the greatest correlation magnitudes were found with Illumina, followed by long amplicon PacBio.
The least spatial structure was detected with PacBio short amplicon sequencing.

The Bray-Curtis metric showed significant ($p < 0.05$) positive correlation when resampling at the same locations one year later (i.e., spatial distance of 0\ m, time lag of 1\ year), for both the total fungal and ECM fungal communities in the long amplicon library.
For the short amplicon library, although the general profile of the correlograms was similar, correlation at 0\ m and 1\ year was not significant, but there was a negative correlation at time lag of 1\ year and a distance of 1\ m for both sequencing technologies.
This puzzling negative correlation was significant in all correlograms based on short amplicon sequencing irrespective of technology.

In contrast to the Bray-Curtis distance, the weighted UNIFRAC distance showed very little spatial structure, with only the total fungal community in the 1\ m 
distance class showing a significant correlation at $p < 0.05$.
No temporal correlation was found for the weighted UNIFRAC distance.

The best fit spatial ranges based on distance-decay curves vary between the different datasets by a factor of about 3, but there is overlap of the 95% confidence intervals for all of the Bray-Curtis spatial ranged in both the total fungal and ECM fungal communities, across amplicon libraries and sequencing technologies (Figure\ \@ref(fig:spatio), Table\ \@ref(tab:variofit)).
Although a distance-decay model was fit for the weighted UNIFRAC distance applied to the total fungal community, the result was very poorly constrained, and a range of 0\ m, indicating no spatial structure, was included in the 95% confidence interval.

@@

```{r variofit, results="asis"}
 rdd(variofit_table) %>%
  mutate_at("conf.low", replace_na, 0) %>%
  mutate(value = ifelse(
    is.na(estimate) | is.na(conf.high),
    "-",
    sprintf("%.2g (%.2g–%.2g)", estimate, conf.low, conf.high)
  )) %>%
  filter(timelag == "0" | term == "timerange") %>%
  pivot_wider(names_from = "term", values_from = "value", id_cols = c("guild", "amplicon", "tech", "metric", "algorithm")) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Algorithm = algorithm, Metric = metric, "Space Range (m)" = range, "Time Range (a)" = timerange) %>%
  arrange(Guild, Amplicon, Tech, Algorithm, Metric) %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)", linesep = c(rep("", 5), "\\addlinespace"))
```

@@

The three sequencing technologies gave similar results for the short amplicon library, the major difference being in sequencing depth.
Although a greater fraction of PacBio raw reads were ultimately mapped to ASVs (`r percent(rdd(bioinf_table)["ITS2","PacBio_Short_reads"]/rdd(rawcounts)["pb_483"], format = "f", digits = 0)`) compared to Illumina (`r percent(rdd(bioinf_table)["ITS2","Illumina_Short_reads"]/rdd(rawcounts)["SH-2257"], format = "f", digits = 0)`) or Ion Torrent (`r percent(rdd(bioinf_table)["ITS2","Ion Torrent_Short_reads"]/rdd(rawcounts)["is_057"], format = "f", digits = 0)`),
the latter two technologies provided much greater sequencing depth for a similar cost, allowing a greater diversity of rare ASVs to be recovered.

DADA2 denoising may perform differently on different technologies (or perhaps sequencing runs), indicated by the fact that clustering ASVs at 97% led to substantially higher correspondence between both the set of sequences recovered from the same library by different technologies (Figure\ \@ref(fig:venn)) and the read counts for each sequence (Figure\ \@ref(fig:read-compare)).
The large number of ASVs unique to Ion Torrent, while only the Illumina dataset recovered an apparent intragenomic variant in the positive control sample, suggests that DADA2 may not control sequencing error as effectively in Ion Torrent sequences as in Illumina, for which it was developed [@callahan2016].

Although the longer read length capabilities of PacBio would allow recovery of longer ITS2 sequences than the other two technologies, PacBio did not recover any ITS2 fragments longer than those recovered by Illumina and Ion Torrent.
Notably, neither long nor short amplicon sequencing recovered any sequences identifiable to *Cantharellus*, an ECM genus which is commonly observed at the study sites as fruitbodies (personal observation), but which is also known to have accelerated evolution in the rDNA [@moncalvo2006] and longer ITS regions than other fungi [@Feibelman_1994], making it an especially difficult target for metabarcoding.
Contrary to expectations, Illumina showed a slightly higher fraction of longer ITS2 sequences than Ion Torrent, which in turn showed slightly longer sequences than PacBio (Supplementary Figures\ \@ref(fig:full-length) and \@ref(fig:ITS2-length)).

Of long amplicon reads, `r percent(rdd(venn_ASV)["1000", "frac_pb_500"])` belonged to ASVs which occurred only in the long amplicon dataset, and clustering at 97% similarity only reduced this fraction to `r percent(rdd(venn_OTU)["1000", "frac_pb_500"])`.
Additionally, ITS2 sequences extracted from the long amplicon dataset included some sequences that were much shorter than those recovered from the short amplicon datasets \@ref(fig:ITS2-length).
Taxonomic assignments revealed that many of these non-shared sequences fall outside kingdom Fungi (Supplementary Figure\ \@ref(fig:heattree-length-compare)), and that in particular the short ITS2 sequences are mostly Alveolates (Supplementary Figure\ \@ref(fig:short-its)).
Within Fungi, the short amplicon datasets recovered more Sordariomycetes (Figure \@ref(fig:pcoa)), possibly because this group was represented by diverse but rare taxa which remained undetected in the PacBio datasets due to the lower sequencing depth.
Additionally, several groups showed anomalous detection in either the long or short datasets, such as *Tulasnella* in the long amplicon dataset (Supplementary Figure \@ref(fig:ecm-heattree)), and *Xylaria* and *Myerozyma* in the short amplicon datasets (Supplementary Figure\ \@ref(fig:heattree-length-compare)).
These differences may be due to primer mismatches in these groups.

@@

ASV recovery for long amplicons using DADA2 was dramatically improved (`r percent(rdd(region_table)$reads[rdd(region_table)$region == "long"]/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` to `r percent((rdd(asvcounts) %>% filter(seq_run == "pb_500", step == "long") %>% pull(reads))/as.integer(gsub(" ", "", rdd(demuxcounts)["pb_500"])), format = "d")` of reads) by denoising homologous subregions independently using the `lsux` and `tzara` packages.
Although newer sequencing platforms from PacBio (Sequel and Sequel II) feature increased sequencing depth and lower error rate compared to the RS II, long sequences inherently require much more sampling depth to identify ASVs.
Thus, `tzara` should increase ASV recovery from these platforms as well.
It may also be adaptable to Oxford Nanopore sequencing, which has hitherto posed difficulties for application to complex community metabarcoding [@loit2019].

@@

The RDP fungal training set and Unite performed comparably at taxonomic placement of long amplicon sequences.
The Warcup database placed notably fewer sequences at all taxonomic levels for all datasets, probably in part due to the fact that only fungal sequences are included.
However, even with this considered, IDTAXA performed very poorly with the Warcup database, placing <25% of ASVs to kingdom in all datasets.
IDTAXA placed fewer sequences than RDPC or SINTAX even with the other databases, but this is expected given its more conservative assignment of confidence scores [@murali2018a].

@@

For ASVs where a long amplicon sequence is available, PHYLOTAX uses phylogenetic relationships to resolve these disagreements in a principled manner.
For instance, `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "Consensus", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()` of Illumina reads were assigned to genus and family, respectively, by the strict consensus of methods, but PHYLOTAX increased this fraction to `r filter(rdd(tax_chart), tech == "Illumina", Algorithm == "PHYLOTAX", rank %in% c("genus", "family"), type == "Reads", reference == "Unite") %>% pull(frac) %>% percent(format = "f", digits = 0) %>% rev() %>% text_list()`.
This led to a corresponding increase in the fraction of reads assigned to a functional guild (Supplementary Figure\ \@ref(fig:ecm)).
